Integrating brain and biomechanical models - a new paradigm for understanding neuro-muscular control

### 2017-11-09 papers/integration/omm_integ.tex
Responds to the comment from R1 about widening projections

\usepackage{url,hyperref,lineno,microtype}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage[onehalfspacing]{setspace}
\usepackage{gensymb} % For \degree
\usepackage[usenames, dvipsnames]{color}

@@

This led us to hypothesize that the retinotopic mapping be accompanied
by an associated widening projection field such that the hill of
activity in superior colliculus is invariant with position on the
collicular surface. There are a number of locations in the system in
which this widening projection field could exist. It could be
implemented in the projections between the retinal populations and the
superficial layer of SC along with the projection between the World
and the FEF population. However, this would affect activity within the
basal ganglia of the model, contradicting a result
in \cite{cope_basal_2017} which explains the `hockey stick' profile
for saccade latency as a function of saccade eccentricity. Instead, we
suggest that a widening projection field is encoded within the
superior colliculus itself, a complex, multi-layered structure which
could quite plausibly support such a function. Indeed, such widening
activity can be seen in the stimulation experiments
in \cite{vokoun_intralaminar_2010}
and \cite{vokoun_response_2014}. Although in this work we do not model
the SC in detail, we extended the model with a third functional layer
named SC\_deep2, shown in Fig.~\ref{scdeep} (\ccg~has only the two
layers SC\_sup and SC\_deep). We introduced a widening projection
based on a Gaussian projection field whose width, $\sigma(r)$ varies
in inverse proportion to the magnification factor, $M(r)$, given in
Eq.~\ref{eq:cmf} according to:
\begin{equation} \label{eq:sigmar}
\sigma(r) = \frac{m_{\sigma}}{M(r)} - \frac{m_{\sigma}}{M^0} + \sigma_0 \qquad r > r_0
\end{equation}
$m_{\sigma}$ is a scalar parameter which determines the `magnitude of
the widening'. $M^0$ is the `starting' magnification factor; within
the foveal region ($0 \leq r \leq r_0$), the projection field is not
allowed to widen and so
\begin{equation} \label{eq:sigmar2}
\sigma(r) = \sigma_0 \qquad r \leq r_0
\end{equation}
which makes $\sigma_0$ the width of the Gaussian projection field
within the foveal region. (Note that the value chosen for the width of
the foveal region, $r_0$ is not identical to the foveal shift
parameter used in the \e{DecayingAtFovea} projections into striatum.)
The \e{Widening Gaussian} projection weight, $w(r,d)$ is then computed
as:
\begin{equation} \label{eq:widening}
w(r,d) = e^{-\frac{d^2}{2\sigma\left(r\right)^2}}
\end{equation}
where $d$ is the distance between the source and destination elements
in the collicular plane.  $m_\sigma$ was set to 50, $\sigma_0$ was
0.3, $M^0$ was 12.43 and $r_0$ was 20.
% Could have a figure here, made from widening_gaussian.m

@@

The saccadic burst generator model was originally conceived with the
assumption of a step input, which returns to zero activity at a
suitable time to curtail the saccade and avoid staircase
saccades \citep{gancarz_neural_1998}. In our model there is no such
mechanism to reduce activity in SC\_deep, and elsewhere.  Although a
successful, accurate saccade towards a target luminance will remove
the excitation which caused the activity in SC\_deep by bringing the
target luminance within the masked, foveal region, the activity in SC
decays too slowly to avoid additional saccadic movements. We found it
necessary to hypothesize an inhibitory feedback mechanism from the SBG
to the brain model.  This is shown in Fig.~\ref{sbg}, which indicates
how the output from the inhibitory burst neurons (IBN) of the SBG
model are used to feed back an inhibitory signal to the SC\_deep,
thalamus and FEF populations in the brain model, resetting them ready
for the next saccade.
% Might have to justify the above a bit more.

@@

To bring our model in line with these results, whilst maintaining the
eccentricity dependent activity in basal ganglia, we hypothesised that
a `widening projection' exists between two maps in superior
colliculus. Activities in one SC\_deep layer remains
eccentricity-dependent, with loops back to thalamus and cortex and
through basal ganglia.  This activity is then fed through a
projection, which applies a Gaussian projection field, whose width
increases with increasing stimulus eccentricity according to
Eq.~\ref{eq:widening}.  The activity in this second SC\_deep layer is
then fed to the weight maps of the SBG. This model was called
`TModel4'. TModel4 was parameterised such that its horizontal and
vertical error was similar---so that its equivalent of
Fig.~\ref{sacc_vs_targ} showed a similar sum of squares error.

>>>>

\usepackage{url,hyperref,lineno,microtype}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage[onehalfspacing]{setspace}
\usepackage{gensymb} % For \degree

@@

This led us to hypothesise that the retinotopic mapping be accompanied
by an associated widening projection field such that the hill of
activity in superior colliculus is invariant with position on the
collicular surface. There are a number of locations in the system in
which this widening projection field could exist. It could be
implemented in the projections between the retinal populations and the
superficial layer of SC along with the projection between the World
and the FEF population. However, this would affect activity within the
basal ganglia of the model, contradicting a result
in \cite{cope_basal_2017} which explains the `hockey stick' profile
for saccade latency as a function of saccade eccentricity. Instead, we
suggest that a widening projection field is encoded within the
superior colliculus itself, a complex, multi-layered structure which
could quite plausibly support such a function. Indeed, such widening
activity can be seen in the stimulation experiments
in \cite{vokoun_intralaminar_2010} and \cite{vokoun_response_2014}.
%
\rvw{In \cite{ghitani_excitatory_2014}, from the same research group,
evidence is presented for an excitatory and widely projecting pathway
from the stratum griseum intermediale (equivalent to our SC\_deep) to
the more superficial layers stratum opticum and stratum griseum
superficiale. Although this pathway is a `wide' projecting field, the
experiments do not indicate whether the projection \e{widens} along
the rostral-caudal axis of the SC. \cite{bayguinov_hard-wired_2015}
presents evidence for another projecting field within SC whose
connectivity pattern \e{does} change along the rostral-caudal
axis. This projection is inhibitory in nature. Although neither of
these results precisely match the widening, excitatory projection
field hypothesised here, they do indicate that such connection
patterns are plausible.}
%
Although in this work we do not model
the SC in detail, we extended the model with a third functional layer
named SC\_deep2, shown in Fig.~\ref{scdeep} (\ccg~has only the two
layers SC\_sup and SC\_deep). We introduced a widening projection
based on a Gaussian projection field whose width, $\sigma(r)$ varies
in inverse proportion to the magnification factor, $M(r)$, given in
Eq.~\ref{eq:cmf} according to:
\begin{equation} \label{eq:sigmar}
\sigma(r) = \frac{m_{\sigma}}{M(r)} - \frac{m_{\sigma}}{M^0} + \sigma_0 \qquad r > r_0
\end{equation}
$m_{\sigma}$ is a scalar parameter which determines the `magnitude of
the widening'. $M^0$ is the `starting' magnification factor; within
the foveal region ($0 \leq r \leq r_0$), the projection field is not
allowed to widen and so
\begin{equation} \label{eq:sigmar2}
\sigma(r) = \sigma_0 \qquad r \leq r_0
\end{equation}
which makes $\sigma_0$ the width of the Gaussian projection field
within the foveal region. (Note that the value chosen for the width of
the foveal region, $r_0$ is not identical to the foveal shift
parameter used in the \e{DecayingAtFovea} projections into striatum.)
The \e{Widening Gaussian} projection weight, $w(r,d)$ is then computed
as:
\begin{equation} \label{eq:widening}
w(r,d) = e^{-\frac{d^2}{2\sigma\left(r\right)^2}}
\end{equation}
where $d$ is the distance between the source and destination elements
in the collicular plane.  $m_\sigma$ was set to 50, $\sigma_0$ was
0.3, $M^0$ was 12.43 and $r_0$ was 20.
% Could have a figure here, made from widening_gaussian.m

@@

The saccadic burst generator model was originally conceived with the
assumption of a step input, which returns to zero activity at a
suitable time to curtail the saccade and avoid staircase
saccades \citep{gancarz_neural_1998}. In our model there is no such
mechanism to reduce activity in SC\_deep, and elsewhere.  Although a
successful, accurate saccade towards a target luminance will remove
the excitation which caused the activity in SC\_deep by bringing the
target luminance within the masked, foveal region, the activity in SC
decays too slowly to avoid additional saccadic movements. We found it
necessary to hypothesise an inhibitory feedback mechanism from the SBG
to the brain model.  This is shown in Fig.~\ref{sbg}, which indicates
how the output from the inhibitory burst neurons (IBN) of the SBG
model are used to feed back an inhibitory signal to the SC\_deep,
thalamus and FEF populations in the brain model, resetting them ready
for the next saccade.
% Might have to justify the above a bit more.

@@

To bring our model in line with these results, whilst maintaining the
eccentricity dependent activity in basal ganglia, we hypothesised that
a `widening projection' exists between two maps in superior
colliculus. \rvw{As described in the Methods section, there is
now experimental evidence for similar projections
\citep{ghitani_excitatory_2014,bayguinov_hard-wired_2015} making this
a plausible suggestion.}  Activities in one SC\_deep layer remains
eccentricity-dependent, with loops back to thalamus and cortex and
through basal ganglia.  This activity is then fed through a
projection, which applies a Gaussian projection field, whose width
increases with increasing stimulus eccentricity according to
Eq.~\ref{eq:widening}.  The activity in this second SC\_deep layer is
then fed to the weight maps of the SBG. This model was called
`TModel4'. TModel4 was parameterised such that its horizontal and
vertical error was similar---so that its equivalent of
Fig.~\ref{sacc_vs_targ} showed a similar sum of squares error.
